---
title: "Mini-Project 3"
author: "Annabel Yim, Kiara Correa Acosta"
date: "April 16, 2019"
output:
  html_document:
    code_folding: hide
---

## To be or not to be Challenged?

Problem Statement: For this assignment, we are trying to analyze and classify the MacLeish trails in terms of difficulty. We will figure this out based on the length and elevation of each trail and categorize them as "Easy", "Moderate" or "Difficult."

```{r, message = FALSE, warning = FALSE, include=FALSE, echo = TRUE}
library(tidyverse)
library(macleish)
library(sf)
```

```{r, message = FALSE, warning = FALSE, echo = TRUE}
trails_data <- macleish_layers %>%
  pluck("trails") %>%
  mutate(computed_length = st_length(geometry))

elevation_data <-macleish_layers %>%
  pluck("contours_3m") 
```
trails_data is a useful data frame created to see the trails layer of the MacLeish package. With this data frame, we are able to see the length and geometry of each trail. elevation_data is also a useful data frame created to see the elevation at each geometry. 

```{r, message = FALSE, warning = FALSE, echo = TRUE}
trails_elevation <- macleish_layers %>%
  pluck("trails") %>%
  mutate(computed_length = st_length(geometry)) %>%
  st_intersection(pluck(macleish_layers ,"contours_3m")) %>%
  group_by(name, computed_length) %>%
  summarize(elevation_gain = max(ELEV_FT) - min(ELEV_FT)) 
```
The trails_elevation data frame was created to see the intersection between the trails and elevation geometries in order to determine the elevations for each trail. Thus, function st_intersection() was used to find the intersections. Because elevation gain is a factor in determining the difficulty of each trail, in order to find the elevation_gains, the trails were first grouped by their names, and the elevation_gain was summarized by taking the summarize() function by subtracting the highest(max) elevation to the lowest(min) elevation to compute the elevation_gains. 

```{r, message = FALSE, warning = FALSE, echo = TRUE}
final_trails_elevation <- trails_elevation %>%
  group_by(name) %>%
  summarize(total_length = sum(computed_length), total_elevation_gain = sum(elevation_gain))
```
In the trails_elevation data frame, there are multiple trails such as the Eastern loop and Western loop that are seen multiple times in the trails_elevation data frame due to the directions in which people can follow the trails. We had to find the total_length and total_elevation gain of these by grouping the trails by the names and then summarizing the total lengths and total elevation gains. And we called this data frame final_trails_elevation.

```{r, message = FALSE, warning = FALSE, echo = TRUE}
trail_names <- c("Eastern Loop", "Western Loop", "Poplar Hill Road", "Porcupine Trail", "Vernal Pool Loop", "entry trail")

trails_difficulty <- final_trails_elevation %>%
  group_by(name, total_elevation_gain, total_length) %>%
  summarize(Shenandoahs_hiking_difficulty = total_elevation_gain*total_length) %>%
  mutate(Shenandoahs_hiking_difficulty = as.numeric(Shenandoahs_hiking_difficulty)) %>%
  arrange(desc(Shenandoahs_hiking_difficulty)) %>%
  filter(name %in% trail_names) 
```
Now that we have the final data frame that can be used to compute the difficulties, we decided to look into existing trail rating systems. One rating system we found interesting was Shenandoah’s hiking difficulty rating, which is the square root of (Elevation gain*2*distance(in miles)). Then they categorized an “easy” trail as being less than 50, moderate as 50-100, and so on. However, because Shenandoah is a national park that is much bigger than MacLeish, we knew that we couldn’t use the same numeric rating. However, we created our own numeric rating by multiplying the elevation gain by distance because what the original formula really does is taking account of these two factors. Then we will determine the difficulties by comparing the relative numeric values among the MacLeish trails. 

Thus, we created the trails_difficulty data frame that computes the difficulty numeric rating by using the summarize() function. Because the total length is in units, we also dropped the units by using as.numeric() for the difficulty ratings. We also realized that the data frame included pathways such as Easy Out and Driveway, that are not trails, so we only filtered by the trails that we are interested in. Lastly, we filtered out the Snowmobile trail because this is a trail that humans do not normally hike on.

```{r, message = FALSE, warning = FALSE, echo = TRUE}
ggplot(trails_difficulty, aes(x = name, y = Shenandoahs_hiking_difficulty)) +
  geom_bar( stat="identity", color = "#008000", fill = "#008000") + labs(x = "trail names", y = "difficulty (hundred thousands)", title = "Difficulties of Trails") +scale_y_continuous(breaks = c(0e+00, 1e+05, 2e+05, 3e+05), label = c("0", "1", "2", "3")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Lastly, after we computed the trail difficulty rating that we had created based on the Shenandoah's difficulty rating system, we decided to draw a bar graph to compare the difficulties. When we look at the bar graph, because Eastern Loop and Western Loop had higher difficulty scales, we decided to categorize them as “Difficult.” Because Poplar Hill Road and Porcupine Trails were in the mid-range, we also decided to categorize them as “Moderate, and Entry Trail and Vernal Pool Loop as "Easy" because they have the smallest difficulty ratings. In the end, using the Existing trails shapefile and 3m contour elevation data, we were able to successfully solve our problem statement of categorizing each MacLeish trails as "Easy", "Moderate" or "Difficult."

One actionable recommendation that Smith can do with the information collected about the difficulty of the trails is to publish this data. Having each trail identifying its level of difficulty would be useful for Smith in the analysis of the trails on the MacLeish property. Our findings can help Smith categorize the trails for people to decide which trail might be best for them. Also, publishing this data to the public would not be a financial burden. The publication of this information would be useful but does not need further accommodations other than making the information accessible to visitors. This project will not take as long and will not have any environmental impacts because the information is already made. One of the only things to take into consideration would be the way in which this data will be published. Some ideas would be having this information on the trail maps and having this on the MacLeish website. In our opinion, there are no roadblocks due to the information already being made. 


```{r, message = FALSE, warning = FALSE, echo = TRUE}
wordcountaddin:::text_stats()
```

^[[Link to the GitHub Repository](https://github.com/annabelyim/sds192-mp3.git]


